Introduction to Course Project
==============================
First load and clean data as per assignment need (remove NA and adjust format)
```{r echo=TRUE}
setwd("~/")
activity <- unzip("activity.zip")
data <- read.csv("activity.csv")
#adjust the date format 
data$date <- as.Date(data$date)
#remove NA from the data set
data <- na.omit(data)

```

Q:1 Histogram of total number of steps taken each day
```{r echo=TRUE}
#count the total number of steps taken per day
Stepsperday <- data %>% group_by(date) %>% summarise(TotalSteps = sum(steps))

#make a histogram of total number of steps taken each day
hist(Stepsperday$TotalSteps, xlab = "Total number of Steps per day", ylab = "Step Counts", main = "Total number of Steps taken each day", col =6)

```

Q:2 Mean and Median number of steps taken each day
```{r echo=TRUE}
#count the mean steps per day
meansteps <- mean(Stepsperday$TotalSteps)
print(meansteps)

#count median steps per day
mediansteps <- median(Stepsperday$TotalSteps)
print(mediansteps)
```

Q:3 Time series plot of the average number of steps taken
```{r echo=TRUE}
#mean steps taken per interval
five_average <- data %>% group_by(interval) %>% summarise(AverageSteps = mean(steps))

#Make time series plot
plot(five_average$interval, five_average$AverageSteps, type = "l", xlab = "Interval", ylab = "Average steps taken during 5 minute interval")
```

Q:4 The 5-minute interval that, on average, contains the maximum number of steps
```{r echo=TRUE}
#calculate the maximum number of steps taken in 5-minute interval
maxstep_interval <- five_average$interval[which.max(five_average$AverageSteps)]
```

Q:5 COde to describe and show a strategy for imputing missing data
```{r echo=TRUE}
#find the missing values and replace it with average value
filldata <- data
for (i in 1:nrow(filldata)) {
  if (is.na(filldata$steps[i])){
    index <- which(filldata$interval[i] == five_average$interval)
    filldata$steps[i] <- five_average[index, ]$AverageSteps
  }
  
}
filldata$date <- as.Date(filldata$date)
```

Q:6 Histogram of the total number of steps taken each day after missing values are imputed
```{r echo=TRUE}
fillstepsperday <- filldata %>% group_by(date) %>% summarise(TotalSteps = sum(steps))
 
#draw histogram for steps taken each day
hist(fillstepsperday$TotalSteps, xlab = "Total number of steps taken each day", ylab = "Count", main = "Histogram of total number of steps taken each day", col =6)
```

Now count the mean total number and median total number of steps per day 

```{r echo=TRUE}
#mean steps per day
fillmeansteps <- mean(fillstepsperday$TotalSteps)

#median steps per day
fillmediansteps <- median(fillstepsperday$TotalSteps)

```

Q:7 Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends.
```{r echo=TRUE}
#first make weekday as variable
filldata$date <- weekdays(filldata$date)

#Now label all days as weekdays
filldata$daytype <- "weekday"

#adjust saturday and sunday as weekends
filldata$daytype[filldata$day %in% c("Saturday", "Sunday")] <- "Weekend"

#Calculate average weekday steps and average weekend steps taken
dayaverage <- filldata %>% group_by(daytype, interval) %>% summarise(AverageSteps = mean(steps))
```
Line plot showing average number of steps taken on weekdays vs on weekends 

```{r echo=TRUE}
library(ggplot2)
qplot(interval, AverageSteps, data = dayaverage, type = "L", geom = "line", xlab = "Interval", ylab = "Average Number of Steps taken", main = "Average steps taken on Weekends vs weekdays", facets = daytype ~.,)

```

